---
title: "Effects plots for *migration and deforestation*, RMR,RK,JPGJ"
date: "27 July 2017"
latest revision: "27_11_2017"
output:
  html_document: default
  word_document: default
authors: Ruth Kelly
---

This RMarkdown document contains code for effects plots in: "Human migration to the forest frontier: implications for conservation management in Madagascar", Jones et al. 2017. 

Required datafiles = 
"data_for_amount_of_deforestation_model.csv"
"data_for_binomial_deforestation_model.csv"
"data_for_migrant_differences_model.csv"
"model_average_2delta_agri_survey_20_07_2017"  
"model_average_2delta_deforestation_binomial" 
"model_average_2delta_diff_migrants"


Note:  Minor editing of figures, e.g. fonts was conducted in a pdf editor after creating these plots. 

```{r setup, include=FALSE, echo=TRUE}


library("reshape2")
library("dplyr")
library("ggplot2")
library("lme4")
library("MuMIn")
library("glmmADMB")
library("grid")

```

#### Figure 4, differences between migrants and non-migrants (Non-economic)

```{r Figure 4 GLMMs differences between migrants and non-migrants, include = TRUE, echo = FALSE}

#### read in data with which model was created. 

pov_mig_std <- read.csv("data_for_migrant_differences_model.csv" )

#####  Read in model averaged object : "model_average_2delta_diff_migrants"       

model1 <- readRDS("model_average_2delta_diff_migrants" )



preds1 <- dplyr::select(pov_mig_std, education, z.hh_age, z.dist2park,
                   z.hh_size, PA)

# New predictors: X1 along the range of original data, other
# variables held constant at their means

preds1$z.hh_age <- mean(preds1$z.hh_age)
preds1$z.hh_size <- mean(preds1$z.hh_size)



x1 <- predict(model1, newdata = preds1,
               se.fit = TRUE, type = "link",
                backtransform = TRUE,
                 full = TRUE)

  x1$fit <- 1 - x1$fit
  x1$upr <- x1$fit+x1$se.fit
  x1$lwr <- x1$fit-x1$se.fit
  x1$dist2park <- pov_mig_std$dist2park
  x1$PA <- pov_mig_std$PA
  x1$education <- pov_mig_std$education
  x1$eduPA <- paste(x1$PA,x1$education, sep = "_" )
x1<- as.data.frame(x1)

x1$education <- relevel(x1$education, ref  = "Low")
######################

# pdf("Fig4_26_11_2017_differences_migrant_nonmigrant.pdf", height = 5, width = 3.5)
ggplot(x1, aes(x = dist2park, y = fit, group = eduPA, linetype = education,  colour = PA, fill = PA)) +
scale_colour_manual(values = c("#D55E00", "#0072B2")) +
scale_fill_manual(values = c("#D55E00", "#0072B2")) +
geom_ribbon(aes(ymin = lwr, ymax = upr, linetype = NA), alpha = 0.40) +
geom_line(size = 1.5) +
  ylim(0,1) +
xlab("Distance from forest (m)") +
  ylab("Proportion of households which are migrants") +
  theme_bw() +
  theme(axis.title = element_text(size = 11)) +
   theme(panel.grid = element_blank()) +
   guides(colour = FALSE, fill = FALSE, linetype = FALSE) +
guides(fill = guide_legend("Protected Area"),linetype = guide_legend("Education"))  +
theme(legend.position = "right", legend.text = element_text(size = 4), legend.title = element_text(size = 4), legend.key.height = unit(0.4, "cm"), legend.key.width = unit(0.4, "cm"))

# dev.off()
##############################

plot(Fig4)

########################


```

#### Figure 5.  Predictors of forest clearance

```{r  binomial models of deforestation full dataset, include = TRUE, echo = FALSE}

### pov_mig_other contains predictor variables except wealth axes.

### Read in data that was used to create the model.. 
clear_data2 <- read.csv("data_for_binomial_deforestation_model.csv")

### Read in model averaging object
model2 <- readRDS("model_average_2delta_deforestation_binomial" )

######## plotting effects...

preds2 <- dplyr::select(clear_data2, PA, z.dist2park, z.hh_age, education, z.PC2, MigrantF, z.PC1, z.hh_size) 
# 
# # New predictors: dist2park along the range of original data, other
# # variables held constant at their means
# 

### First plot distance to forest by education and Protected area. 
preds2$z.hh_age <- mean(preds2$z.hh_age)
preds2$z.hh_size <- mean(preds2$z.hh_size)
preds2$z.PC1 <- mean(preds2$z.PC1)
preds2$z.PC2 <- mean(preds2$z.PC2)


 x1 <- predict(model2, newdata = preds2,
                se.fit = TRUE, type = "link",
                 backtransform = TRUE,
                  full = TRUE)


x1$fit <- x1$fit
x1$upr <- x1$fit+x1$se.fit
x1$lwr <- x1$fit-x1$se.fit
x1$dist2park <- clear_data2$dist2park
x1$PA <- clear_data2$PA
x1$education<- clear_data2$education
x1$MigrantF<- clear_data2$MigrantF

x1$eduPAmig <- paste(x1$PA,x1$education, x1$MigrantF, sep = "_" )   
x1$eduPA <- paste(x1$PA,x1$education, sep = "_" )
x1$PAmig <- paste(x1$PA,x1$Migrant, sep = "_" )
x1<- as.data.frame(x1)       

###
 x1$education <- relevel(x1$education, ref  = "Low")      
# ######################
# 

x2 <- x1[x1$education == "Low",]
#################

plot_dist2 <- ggplot(x2, aes(x = dist2park, y = fit, group = eduPAmig, colour = eduPA, fill = eduPA, linetype = MigrantF)) + 
   ylim(0,1.05) + 
   scale_colour_manual(values = c("#D55E00", "#0072B2")) +
scale_fill_manual(values = c("#D55E00", "#0072B2"), labels = c("Established", "New")) +
 geom_ribbon(aes(ymin = lwr, ymax = upr, linetype = NA), alpha = 0.40) + 
 geom_line(size = 1.5) +
 xlab("Distance from forest (m)") + 
   ylab("Probability of having cleared land") +
   theme_bw() + 
   theme(axis.title = element_text(size = 11)) + 
    theme(panel.grid = element_blank()) + 
    guides(colour = FALSE, fill = FALSE, linetype = FALSE) 


#####

####  Same but for household age.. 


preds3 <- dplyr::select(clear_data2, PA, z.dist2park, z.hh_age, education, z.PC2, MigrantF, z.PC1, z.hh_size) 
# 
# # New predictors: household age along the range of original data, other
# # variables held constant at their means
# 

# 
### Second plot household age  by education and Protected area. 
preds3$z.dist2park <- mean(preds2$z.dist2park)
preds3$z.hh_size <- mean(preds2$z.hh_size)
preds3$z.PC1 <- mean(preds2$z.PC1)
preds3$z.PC2 <- mean(preds2$z.PC2)


 x3 <- predict(model2, newdata = preds3,
                se.fit = TRUE, type = "link",
                 backtransform = TRUE,
                  full = TRUE)


x3$fit <- x3$fit
x3$upr <- x3$fit+x3$se.fit
x3$lwr <- x3$fit-x3$se.fit
x3$Household_age <- clear_data2$hh_age
x3$PA <- clear_data2$PA
x3$education<- clear_data2$education
x3$MigrantF<- clear_data2$MigrantF

x3$eduPAmig <- paste(x3$PA, x3$education, x3$MigrantF, sep = "_" )   
x3$eduPA <- paste(x3$PA,x3$education, sep = "_" )
x3$PAmig <- paste(x3$PA,x3$Migrant, sep = "_" )
x3<- as.data.frame(x3)       


 x3$education <- relevel(x3$education, ref  = "Low")      
# ######################
# 

x4 <- x3[x3$education == "Low",]
#################

    
#########

plot_age <- ggplot(x4, aes(x = Household_age, y = fit, group = eduPAmig, colour = eduPA, fill = eduPA, 
linetype = MigrantF)) + 
   scale_colour_manual(values = c("#D55E00", "#0072B2")) +
scale_fill_manual(values = c("#D55E00", "#0072B2"), labels = c("Established", "New")) +
 geom_ribbon(aes(ymin = lwr, ymax = upr, linetype = NA), alpha = 0.40) + 
 geom_line(size = 1.5) +
 xlab("Household age (yrs)") + 
   ylab("") +
  ylim(0,1.05) + 
   theme_bw() + 
   theme(axis.title = element_text(size = 11)) + 
    theme(panel.grid = element_blank()) + 
    guides(colour = FALSE, fill = FALSE, linetype = FALSE) + 
    guides(fill = guide_legend("Protected Area"), linetype = guide_legend("Migration")) +  
  theme(legend.position = "right", legend.text = element_text(size = 9), legend.title = element_text(size = 9)) 
 



plot(plot_dist2)
plot(plot_age)

# ## output as PDF
# pdf("Fig5_deforestation_models.pdf", width = 7.4, height = 4)
# grid.draw(cbind(ggplotGrob(plot_dist2), ggplotGrob(plot_age), size="last"))
# dev.off()

```
  
#### Supplementary Figure A4. Predictors of amount of land cleared based on agricultural data subset. 

```{r models of deforestation agricultural subset, include = TRUE, echo = FALSE}

### read in data used in model construction
land_data2 <- read.csv("data_for_amount_of_deforestation_model.csv")

### read in Model averaging object

model3 <- readRDS("model_average_2delta_agri_survey_20_07_2017")



preds_ag <- dplyr::select(land_data2, education, z.hh_age, z.dist2park,
                   z.PC1, PA)

# New predictors:  variable of interest along original data, other
# variables held constant at their means



preds_ag$z.hh_age <- mean(preds_ag$z.hh_age)
preds_ag$z.PC1 <- mean(preds_ag$z.PC1)

x2_ag <- predict(model3, newdata = preds_ag,
               se.fit = TRUE, type = "link",
               full = TRUE)

######## backtransform need to be conducted here as it was not conducted in the predict function above.

bkfit <- exp(x2_ag$fit)
bkupr <- exp(x2_ag$fit + x2_ag$se.fit)
bklwr <- exp(x2_ag$fit - x2_ag$se.fit)

####

x_ag <- as.data.frame(x2_ag)
 ####  
 
  x_ag$fit <- bkfit/(10^6)
  x_ag$upr <- bkupr/(10^6)
  x_ag$lwr <- bklwr/(10^6)
  x_ag$dist2park <- land_data2$dist2park
  x_ag$PA <- preds_ag$PA
  x_ag$education <- preds_ag$education
  x_ag$Migrant <- land_data2$MigrantF

  x_ag$eduPAmig <- paste(x_ag$PA,x_ag$education, x_ag$MigrantF, sep = "_" )   
x_ag$eduPA <- paste(x_ag$PA,x_ag$education, sep = "_" )
x_ag$PAmig <- paste(x_ag$PA,x_ag$Migrant, sep = "_" )

x_ag$education <- relevel(x_ag$education, ref  = "Low")
######################


x_ag2 <- x_ag[x_ag$education == "Low",]

x_ag2 <- droplevels(x_ag2)
##

x_ag2$PA <- relevel(x_ag2$PA, ref = "Established")

x_ag2$PA <- relevel(x_ag2$PA ,
                ref ="New")


plot_dist_ag <-  ggplot(x_ag2, aes(x = dist2park, y = fit, group = PAmig, colour = PA, fill = PA, 
linetype = Migrant)) + 
ylim(0, 0.1) + 
   scale_colour_manual(values = c("#D55E00", "#0072B2")) +
scale_fill_manual(values = c("#D55E00", "#0072B2"), labels = c("Established", "New")) +
  geom_ribbon(aes(ymin = lwr, ymax = upr, linetype = NA), alpha = 0.4) + 
 geom_line(size = 1.5) +
 xlab("Distance from park (m)") + 
   ylab("Amount of land cleared from forest per household (km2)") +
   theme_bw() + 
   theme(axis.title = element_text(size = 11)) + 
    theme(panel.grid = element_blank()) + 
    guides(colour = FALSE, fill = FALSE, linetype = FALSE) 


#### and again for PC1 !

preds_ag <- dplyr::select(land_data2, education, z.hh_age, z.dist2park,
                   z.PC1, PA)

# New predictors: x variable of interest along original data, other
# variables held constant at their means

preds_ag$z.hh_age <- mean(preds_ag$z.hh_age)
preds_ag$z.dist2park <- mean(preds_ag$z.dist2park)

x2_ag <- predict(model3, newdata = preds_ag,
               se.fit = TRUE, type = "link",
               full = TRUE)

######## backtransform need to be conducted here as it was not conducted in the predict function above.

bkfit <- exp(x2_ag$fit)
bkupr <- exp(x2_ag$fit + x2_ag$se.fit)
bklwr <- exp(x2_ag$fit - x2_ag$se.fit)

####
x_ag<- as.data.frame(x2_ag)

 ####  
  x_ag$fit <- bkfit/(10^6)
  x_ag$upr <- bkupr/(10^6)
  x_ag$lwr <- bklwr/(10^6)
  x_ag$PC1 <- land_data2$PC1
  x_ag$PA <- preds_ag$PA
  x_ag$education <- preds_ag$education
  x_ag$Migrant <- land_data2$MigrantF
x_ag$eduPAmig <- paste(x_ag$PA,x_ag$education, x_ag$MigrantF, sep = "_" )   
x_ag$eduPA <- paste(x_ag$PA,x_ag$education, sep = "_" )
x_ag$PAmig <- paste(x_ag$PA,x_ag$Migrant, sep = "_" )

x_ag$education <- relevel(x_ag$education, ref  = "Low")
######################


x_ag2 <- x_ag[x_ag$education == "Low",]
x_ag2 <- droplevels(x_ag2)
##

x_ag2$PA <- relevel(x_ag2$PA ,
                ref ="New")


################  Plot showing PC1


plot_PC1_ag <-  ggplot(x_ag2, aes(x = PC1, y = fit, group = PAmig, colour = PA, fill = PA, 
linetype = Migrant)) + 
   scale_colour_manual(values = c("#D55E00", "#0072B2")) +
scale_fill_manual(values = c("#D55E00", "#0072B2"), labels = c("Established", "New")) +
  geom_ribbon(aes(ymin = lwr, ymax = upr, linetype = NA), alpha = 0.4) + 
 geom_line(size = 1.5)+
 ylim(c(0,1.8)) + 
 xlab("Wealth axis - Low to High - PC1") + 
   ylab("") +
   theme_bw() + 
   theme(axis.title = element_text(size = 11)) + 
    theme(panel.grid = element_blank()) + 
    guides(colour = FALSE, fill = FALSE, linetype = FALSE) + 
  guides(linetype = guide_legend("Migration"),fill = guide_legend("Protected Area")) +  
  theme(legend.position = "right", legend.text = element_text(size = 9), legend.title = element_text(size = 9)) 
  

plot(plot_dist_ag)
plot(plot_PC1_ag)

# ## output as PDF
# pdf("Supp_Fig4_Deforestation_agri_areas3.pdf", width = 7.4, height = 4.2)
# grid.draw(cbind(ggplotGrob(plot_dist_ag), ggplotGrob(plot_PC1_ag), size="last"))
# dev.off()

###

```
